---
title: "Montreal BCM Analysis"
author: "Caitlin Ward"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 14, fig.height = 6,
                      warning = FALSE)

library(ggplot2)
library(grid)
library(gridExtra)
library(nimble)
library(plyr)
library(knitr)
library(scales)
library(kableExtra)

# functions to calculate the alarms
source('../scripts/modelCodes.R')

theme_set(theme_bw() + 
            theme(strip.background = element_rect(fill = 'white'),
                  strip.text = element_text(size = 18),
                  axis.title = element_text(size = 18),
                  axis.text = element_text(size = 16),
                  legend.title = element_text(size = 18),
                  legend.text = element_text(size = 16)))

infPeriodSpec <- 'exp'

```


## Data

```{r}
### read data
montreal <- read.csv('./Data/montrealClean.csv')
montreal$date <- as.Date(montreal$date)

ggplot(montreal, aes(x = date, y = smoothedCases, col = factor(peak))) +
  geom_point() + 
  labs(col = "Peak", x = "Date", y = "Incidence")

```


## Gelman Rubin

Number of models remaining with Gelman Rubin > 1.1.

```{r}
grAll <- readRDS('./resultsFinal/grAll.rds')
grAll <- grAll[grAll$infPeriod == infPeriodSpec,]

# which didn't converge
notConverge <- grAll[which(grAll$gr > 1.1),  ]
notConvergeModels <-  notConverge[
  !duplicated(notConverge
              [,-which(colnames(notConverge) %in% c('gr', 'grUpper', 'param'))]),
  c('alarmFit', 'infPeriod', 'peak')]
notConvergeModels$noConverge <- 1


table(notConvergeModels$alarmFit,
      notConvergeModels$peak)

```


## Results by peak

```{r}
################################################################################
### load posterior estimates of alarm functions
alarmAll <- readRDS('./resultsFinal/alarmPostAll.rds')
alarmAll <- alarmAll[alarmAll$infPeriod == infPeriodSpec,]

### remove those that did not converge (TEMPORARY)
alarmAll <- merge(alarmAll, notConvergeModels,
                  by = c('alarmFit', 'infPeriod', 'peak'),
                  all.x = T)
alarmAll$noConverge[is.na(alarmAll$noConverge)] <- 0

# format for better plotting
alarmAll$alarmFit <- factor(alarmAll$alarmFit,
                            levels = c('spline', 'gp', 'thresh', 'hill', 'power'),
                            labels = c('Spline', 'Gaussian Process', 
                                       'Threshold', 'Hill', 'Power'))


################################################################################
### load posterior estimates of R0 over epidemic time
r0All <- readRDS('./resultsFinal/r0PostAll.rds')
r0All <- r0All[r0All$infPeriod == infPeriodSpec,]

### remove those that did not converge (TEMPORARY)
r0All <- merge(r0All, notConvergeModels,
               by = c('alarmFit', 'infPeriod', 'peak'),
               all.x = T)
r0All$noConverge[is.na(r0All$noConverge)] <- 0

# format for better plotting
r0All$alarmFit <- factor(r0All$alarmFit,
                         levels = c('spline', 'gp', 'thresh', 'hill', 'power',
                                    'betatSpline', 'basic'),
                         labels = c('Spline', 'Gaussian Process', 
                                    'Threshold', 'Hill', 'Power',
                                    'Beta[t]', 'Basic'))
r0All <- r0All[order(r0All$alarmFit, 
                     r0All$peak, 
                     r0All$time),]

################################################################################
# format montreal data to merge with R0 - those always start at "time 1"

# add 100 empty
montrealNew <- montreal

idxStart <- 15

montrealNew$timeFull <- 1:nrow(montrealNew)-idxStart
montrealNew$timeFull[1:idxStart] <- NA
montrealNew$timePeak5 <- montrealNew$timePeak4 <- montrealNew$timePeak3 <- 
    montrealNew$timePeak2 <- montrealNew$timePeak1 <- NA

# peak 1 time 1 is row 6
montrealNew$timePeak1[(idxStart + 1):(max(which(montrealNew$peak == 1)))] <-
  1:(max(which(montrealNew$peak == 1)) - idxStart)

montrealNew$timePeak2[(min(which(montrealNew$peak == 2)) + 1):(max(which(montrealNew$peak == 2)))] <-
  (min(which(montrealNew$peak == 2)) + 1):(max(which(montrealNew$peak == 2))) -
  min(which(montrealNew$peak == 2))

montrealNew$timePeak3[(min(which(montrealNew$peak == 3)) + 1):(max(which(montrealNew$peak == 3)))] <-
  (min(which(montrealNew$peak == 3)) + 1):(max(which(montrealNew$peak == 3))) -
  min(which(montrealNew$peak == 3))

montrealNew$timePeak4[(min(which(montrealNew$peak == 4)) + 1):(max(which(montrealNew$peak == 4)))] <-
  (min(which(montrealNew$peak == 4)) + 1):(max(which(montrealNew$peak == 4))) -
  min(which(montrealNew$peak == 4))

montrealNew$timePeak5[(min(which(montrealNew$peak == 5)) + 1):(max(which(montrealNew$peak == 5)))] <-
  (min(which(montrealNew$peak == 5)) + 1):(max(which(montrealNew$peak == 5))) -
  min(which(montrealNew$peak == 5))

################################################################################
# WAIC

options(scipen=20)
waicAll <- readRDS('./resultsFinal/waicAll.rds')
waicAll <- waicAll[waicAll$infPeriod == infPeriodSpec,]

### remove those that did not converge (TEMPORARY)
waicAll <- merge(waicAll, notConvergeModels,
                 by = c('alarmFit', 'infPeriod', 'peak'),
                 all.x = T)

waicAll$noConverge[is.na(waicAll$noConverge)] <- 0

waicAll <- waicAll[order(waicAll$peak, waicAll$waic),]
# summaries of WAIC values by model
waicAll$waic <- round(waicAll$waic, 2)
waicAll$lppd <- round(waicAll$lppd, 2)
waicAll$pWAIC <- round(waicAll$pWAIC, 2)


```

### Peak 1

#### Data

```{r}
ggplot(subset(montreal, peak== '1'), aes(x = date, y = smoothedCases)) +
  geom_line() + 
  labs(x = "Date", y = "Incidence")
```

#### Estimated Alarm Functions

```{r}
ggplot(subset(alarmAll, peak== '1' & noConverge == 0), aes(x = xAlarm, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = '30-day incidence', y = 'Alarm')
```

#### Estimated R0 Functions

```{r}
r0Peak1 <- subset(r0All, peak == '1' & noConverge == 0)
r0Peak1 <- merge(r0Peak1, montrealNew, 
                 by.x = 'time', by.y = 'timePeak1', 
                 all.x = T)

ggplot(r0Peak1, aes(x = date, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = 'Date', y = 'R0') + 
  geom_hline(yintercept = 1, linetype = 2)
```


#### WAIC

```{r}
waicPeak <- waicAll[waicAll$peak == '1',]
kable(waicPeak[,-which(colnames(waicPeak) %in% c('peak', 'infPeriod'))], 
      row.names = F)  %>% 
  kable_styling()
```


### Peak 2

#### Data

```{r}
ggplot(subset(montreal, peak== '2'), aes(x = date, y = smoothedCases)) +
  geom_line() + 
  labs(x = "Date", y = "Incidence")
```

#### Estimated Alarm Functions

```{r}
ggplot(subset(alarmAll, peak== '2'  & noConverge == 0), aes(x = xAlarm, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = '30-day incidence', y = 'Alarm')
```

#### Estimated R0 Functions

```{r}
r0Peak2 <- subset(r0All, peak == '2' & noConverge == 0)
r0Peak2 <- merge(r0Peak2, montrealNew, 
                 by.x = 'time', by.y = 'timePeak2', 
                 all.x = T)

ggplot(r0Peak2, aes(x = date, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = 'Date', y = 'R0') + 
  ylim(0.5, 1.25) + 
  geom_hline(yintercept = 1, linetype = 2)
```


#### WAIC

```{r}
waicPeak <- waicAll[waicAll$peak == '2',]
kable(waicPeak[,-which(colnames(waicPeak) %in% c('peak', 'infPeriod'))], 
      row.names = F)  %>% 
  kable_styling()
```


### Peak 3

#### Data

```{r}
ggplot(subset(montreal, peak== '3'), aes(x = date, y = smoothedCases)) +
  geom_line() + 
  labs(x = "Date", y = "Incidence")
```

#### Estimated Alarm Functions

```{r}
ggplot(subset(alarmAll, peak== '3' & noConverge == 0), aes(x = xAlarm, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = '30-day incidence', y = 'Alarm')
```

#### Estimated R0 Functions

```{r}
r0Peak3 <- subset(r0All, peak == '3' & noConverge == 0)
r0Peak3 <- merge(r0Peak3, montrealNew, 
                 by.x = 'time', by.y = 'timePeak3', 
                 all.x = T)

ggplot(r0Peak3, aes(x = date, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = 'Date', y = 'R0') + 
  ylim(0.5, 1.6) + 
  geom_hline(yintercept = 1, linetype = 2)
```


#### WAIC

```{r}
waicPeak <- waicAll[waicAll$peak == '3',]
kable(waicPeak[,-which(colnames(waicPeak) %in% c('peak', 'infPeriod'))], 
      row.names = F)  %>% 
  kable_styling()
```


### Peak 4

#### Data

```{r}
ggplot(subset(montreal, peak== '4'), aes(x = date, y = smoothedCases)) +
  geom_line() + 
  labs(x = "Date", y = "Incidence")
```

#### Estimated Alarm Functions

```{r}
ggplot(subset(alarmAll, peak== '4' & noConverge == 0), aes(x = xAlarm, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = '30-day incidence', y = 'Alarm')
```

#### Estimated R0 Functions

```{r}
r0Peak4 <- subset(r0All, peak == '4' & noConverge == 0)
r0Peak4 <- merge(r0Peak4, montrealNew, 
                 by.x = 'time', by.y = 'timePeak4', 
                 all.x = T)

ggplot(r0Peak4, aes(x = date, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = 'Date', y = 'R0') + 
  ylim(0.5, 2) + 
  geom_hline(yintercept = 1, linetype = 2)
```

#### WAIC

```{r}
waicPeak <- waicAll[waicAll$peak == '4',]
kable(waicPeak[,-which(colnames(waicPeak) %in% c('peak', 'infPeriod'))], 
      row.names = F) %>% 
  kable_styling()
```


### Peak 5

#### Data

```{r}
ggplot(subset(montreal, peak== '5'), aes(x = date, y = smoothedCases)) +
  geom_line() + 
  labs(x = "Date", y = "Incidence")
```

#### Estimated Alarm Functions

```{r}
ggplot(subset(alarmAll, peak== '5' & noConverge == 0), aes(x = xAlarm, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = '30-day incidence', y = 'Alarm')
```

#### Estimated R0 Functions

```{r}
r0Peak5 <- subset(r0All, peak == '5' & noConverge == 0)
r0Peak5 <- merge(r0Peak5, montrealNew, 
                 by.x = 'time', by.y = 'timePeak5', 
                 all.x = T)

ggplot(r0Peak5, aes(x = date, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = 'Date', y = 'R0') + 
  ylim(0.5, 2) + 
  geom_hline(yintercept = 1, linetype = 2)
```

#### WAIC

```{r}
waicPeak <- waicAll[waicAll$peak == '5',]
kable(waicPeak[,-which(colnames(waicPeak) %in% c('peak', 'infPeriod'))], 
      row.names = F) %>% 
  kable_styling()
```



### Full

#### Data

```{r}
ggplot(montreal, aes(x = date, y = smoothedCases)) +
  geom_line() + 
  labs(x = "Date", y = "Incidence")
```

#### Estimated Alarm Functions

```{r}
ggplot(subset(alarmAll, peak== 'full' & noConverge == 0), aes(x = xAlarm, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = '30-day incidence', y = 'Alarm')
```

#### Estimated R0 Functions

```{r}
r0PeakFull <- subset(r0All, peak == 'full' & noConverge == 0)
r0PeakFull <- merge(r0PeakFull, montrealNew, 
                    by.x = 'time', by.y = 'timeFull', 
                    all.x = T)

ggplot(r0PeakFull, aes(x = date, y = mean)) +  
  geom_line() +
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.3, fill = 'blue') +
  facet_wrap(~alarmFit, nrow = 1) +
  labs(x = 'Date', y = 'R0') + 
  geom_hline(yintercept = 1, linetype = 2)
```


#### WAIC

```{r}
waicPeak <- waicAll[waicAll$peak == 'full',]
kable(waicPeak[,-which(colnames(waicPeak) %in% c('peak', 'infPeriod'))], 
      row.names = F)  %>% 
  kable_styling()
```


## All together

#### Estimated Alarm Functions

```{r}
ggplot(subset(alarmAll, noConverge == 0), aes(x = xAlarm, y = mean, 
                                              group =  alarmFit)) +  
  geom_line(aes(col = alarmFit)) +
  geom_ribbon(aes(ymin=lower, ymax=upper, fill = alarmFit), alpha=0.3) +
  facet_wrap(~peak, nrow = 2, scales = 'free_x') +
  labs(x = '30-day incidence', y = 'Alarm', col = 'Alarm', fill = 'Alarm') 
```

#### Estimated R0 Functions

```{r}

r0PeakAll <- rbind.fill(r0Peak1, r0Peak2, r0Peak3, r0Peak4, r0PeakFull)

ggplot(r0PeakAll, aes(x = date, y = mean, group = alarmFit)) +  
  geom_line(aes(col = alarmFit)) +
  geom_ribbon(aes(ymin=lower, ymax=upper, fill = alarmFit), alpha=0.3) +
  facet_wrap(~peak.x, nrow = 2, scales = 'free') +
  labs(x = 'Date', y = 'R0', col = 'Alarm', fill = 'Alarm') + 
  geom_hline(yintercept = 1, linetype = 2)
```


#### WAIC 

```{r}
kable(waicAll[,-which(colnames(waicAll) %in% c('peak', 'infPeriod'))], 
      row.names = F) %>% 
  pack_rows(index = table(waicAll$peak))
```




